Texture Mapping 3D Planar Models of Indoor Environments with Noisy Camera Poses

### 2012-12-17 spie/_region_.tex
added spie dir

\documentclass[10pt,twocolumn,letterpaper]{article}

@@

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the 3DIMPVT Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%%%%%%%%% TITLE
\title{Texture Mapping 3D Models of Indoor Environments with Noisy
  Camera Poses}

\author{First Author\\
  Institution1\\
  Institution1 address\\
  {\tt\small firstauthor@i1.org}
  % For a paper whose authors are all at the same institution, omit
  % the following lines up until the closing ``}''.  Additional
  % authors and addresses can be added with ``\and'', just like the
  % second author.  To save space, use either the email address or
  % home page, not both
  \and
  Second Author\\
  Institution2\\
  First line of institution2 address\\
  {\small\url{http://www.author.org/~second}} }

\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuties when
  successive images are projected adjacently onto a plane for
  texturing. We propose two approaches to reduce discontinuities in
  texture mapping 3D models made of planar surfaces. The first one is
  tile based and can be used for images and planes at arbitrary angles
  relative to each other. The second one results in a more seamless
  texture, but is only applicable where camera axes for images are
  closely aligned with plane normals of the surfaces to be
  textured. The effectiveness of our approaches are demonstrated on
  two indoor datasets.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

@@

\begin{figure}[h!]
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_naive.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog_noblend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift_blend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog.jpg}}
  \caption{(a) Direct mapping. (b) Mapping with caching. (c) Mapping
    with caching after image alignment. (d) Seam minimization after
    image alignment). (e) same as (c) with blending. (f) same as (d)
    with blending.}
  \label{fig:compareAll}
\end{figure}

@@

  \centering \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}
  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing}
  \label{fig:mosaic3D}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3.5in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

In this section, we describe our proposed method for seamless texture
mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we project these images onto the plane
and perform rotation and shifting in order to maximize SIFT feature
matches in overlapping areas between these projections. We then select
which textures to be used either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method to be described in Section \ref{sec:seamMinimization}. This
choice depends on the availability of head-on images for a given
plane. We then finish by applying linear alpha blending to smooth out
any remaining seams.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]

If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. To fix this we choose the first image
to serve as the anchor for the rest, meaning all the absolute
distances are based on its original location. This is done by adding a
row with a $1$ in the first column and the rest zeros.

@@

For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the scoring function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus, the scoring function is less
signifcant for walls, and it is conceivable to use a different
texturing strategy for them so as to minimize visible seams in texture
mapped planes. This can be done by choosing the smallest possible set
of images that (a) covers the entire plane and (b) minimizes the
amount of visible seams between them. A straigntforward cost function
that accomplishes the latter is the Sum of Squared Differences (SSD)
of pixels in overlapping regions between all pairs of images, after
they have been rotated as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the plane, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images.  This is a
complex problem, though we can take a number of steps to simplify
it. Given that wall-texture candidate images are taken from more or
less head-on angles, and assuming only minor rotations are made in
Section \ref{sec:projectionAndRotation}, we can reason that their
projections onto the plane are approximately rectangular. By cropping
them all to be rectangular, as shown in Figure
\ref{fig:projectionAngles}(c), our problem becomes the conceptually
simpler one of filling a polygon with rectangles, such that the sum of
all costs between each pair of rectangles is minimal. We thus also
retain the advantages of working with rectangular units, as explained
in Section \ref{sec:simpleTextureMapping}.

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process.}
  \label{fig:dagCreation}
\end{figure}

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images\footnote{ For the seam minimization
  process, it is important that images only contain content that
  should be mapped onto the target plane in question. The tiling
  approach checks occlusion for each tile as it is being textured,
  whereas the seam minimization approach uses entire images.
  Occlusion checks thus need to be performed over the entirety of each
  image to determine available areas for texture mapping a given
  plane.  }.

@@

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size. For the seam minimization method, we have
already ensured overlap between images. To enforce consistent blending
however, we add a minimuum required overlap distance while solving the
shortest path problem in Section
\ref{sec:seamMinimization}. Additionally, if images overlap in a
region greater than the overlap distance, we only apply blending over
an area equal to the overlap distance.

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in]{4thfloor8.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{floorcropped.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) a ceiling, (c) an entire floor, (d) a full model.}
  \label{fig:results}
\end{figure}

@@

>>>>

\message{ !name(paper.tex)}\documentclass[10pt,twocolumn,letterpaper]{article}

@@

 \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{84} % *** Enter the 3DIMPVT Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

\message{ !name(paper.tex) !offset(-3) }

%%%%%%%%% TITLE
\title{Texture Mapping 3D Planar Models of Indoor Environments with Noisy
  Camera Poses}

\author{Peter Cheng\\
  UC Berkeley
  \and
  Michael Anderson
  \and
  Stewart He
  \and
  Avideh Zakhor
  %Institution1 address\\
  %{\tt\small firstauthor@i1.org}
  % For a paper whose authors are all at the same institution, omit
  % the following lines up until the closing ``}''.  Additional
  % authors and addresses can be added with ``\and'', just like the
  % second author.  To save space, use either the email address or
  % home page, not both
  %\and
  %Second Author\\
  %\and
  %Third Author\\
  %\and
  %Fourth Author\\
%  Institution2\\
%  First line of institution2 address\\
%  {\small\url{http://www.author.org/~second}} }
}
\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuities when
  successive images are projected adjacently onto a plane for
  texturing. We propose two approaches to reduce discontinuities in
  texture mapping 3D models made of planar surfaces. The first one is
  tile based and can be used for images and planes at arbitrary angles
  relative to each other. The second one results in a more seamless
  texture, but is only applicable where camera axes for images are
  closely aligned with plane normals of the surfaces to be
  textured. The effectiveness of our approaches are demonstrated on
  two indoor datasets.
\end{abstract}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

@@

\begin{figure}[h!]
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_naive.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_cache.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_cache_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_dynprog_noblend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_cache_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_dynprog.pdf}}
  \caption{(a) Direct mapping. (b) Mapping with caching. (c) Mapping
    with caching after image alignment. (d) Seam minimization after
    image alignment). (e) same as (c) with blending. (f) same as (d)
    with blending.}
  \label{fig:compareAll}
\end{figure}

@@

  \centering \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}
  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing.}
  \label{fig:mosaic3D}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

In this section, we describe our proposed method for seamless texture
mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we project all images onto the plane and
rotate them for verticality. Next,we perform 2D shifting in order to
maximize SIFT feature matches in overlapping areas between these
projections. We then select which textures to be used either with the
tile caching method from Section \ref{sec:mappingWithCaching}, or with
the more specialized method to be described in Section
\ref{sec:seamMinimization}. This choice depends on the availability of
head-on images for a given plane. We then finish by applying linear
alpha blending to smooth out any remaining seams.

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. To fix this we choose the first image
to serve as the anchor for the rest, meaning all the absolute
distances are based on its original location. This is done by adding a
row with a $1$ in the first column and the rest zeros.

@@

For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the scoring function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus, the scoring function is less
significant for walls, and it is conceivable to use a different
texturing strategy for them so as to minimize visible seams in texture
mapped planes. This can be done by choosing the smallest possible set
of images that (a) covers the entire plane and (b) minimizes the
amount of visible seams between them. A straightforward cost function
that accomplishes the latter is the Sum of Squared Differences (SSD)
of pixels in overlapping regions between all pairs of images, after
they have been rotated as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the plane, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images.  This is a
complex problem, though we can take a number of steps to simplify it.

Given that wall-texture candidate images are taken from more or less
head-on angles, and assuming only minor rotations are made in Section
\ref{sec:projectionAndRotation}, we can reason that their projections
onto the plane are approximately rectangular. By cropping them all to
be rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:simpleTextureMapping}.

@@

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

@@

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the
posters in the images, which have clear misalignment over seams in
Figure \ref{fig:compareAll}(c), but are much more aligned in Figure
\ref{fig:compareAll}(d). This is because the seam minimization
approach directly reduces the cost of each image boundary, while the
tile caching method uses a scoring function that only approximates
this effect. Furthermore, seam minimization guarantees the best
selection of images, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. The seam minimization approach
is also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images\footnote {For the seam minimization
  approach, occlusion checks are performed over the entirety of each
  image to determine available areas for texture mapping. Since indoor
  environments mostly consist of vertical or horizontal surfaces with
  high amounts of right angles, image projections remain largely
  rectangular after masking out occluded areas. Therefore, as a
  preprocessing step, we can recursively split each image into
  rectangular pieces and check for occlusions in a similar manner as
  in the tiling process.  }.

@@

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size. For the seam minimization method, we have
already ensured overlap between images. To enforce consistent blending
however, we add a minimum required overlap distance while solving the
shortest path problem in Section
\ref{sec:seamMinimization}. Additionally, if images overlap in a
region greater than the overlap distance, we only apply blending over
an area equal to the overlap distance.

@@

\begin{figure*}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}}
  ~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering
  \subfloat[][]{\includegraphics[height=2in, width=3in]{floorcropped.jpg}}
  ~~~~~~~~
  \centering \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) an entire floor, (d) a full model.}
  \label{fig:results}
\end{figure*}

@@

\message{ !name(paper.tex) !offset(-5) }
